### 2023-03-06 Manuscript/Draft1.Rmd
update manuscript

```{r strata-plot, echo=FALSE,out.width="95%",fig.show='hold', fig.align='center', fig.cap="Strata inhabited by each species in our population models."}
knitr::include_graphics(c("Images/Strata.png"))
knitr::include_graphics(c("Images/Excluded_strata.png"))
```

@@

@@

Figure \@ref(fig:PopPct) depicts the spatial shifting that occurs in our population models in each of the survey weeks.

Of the scenarios without covariates, $33/48\approx68$\% had a VAST fit with a lower relative error compared to the stratified mean estimate. All 15 of the covariate-free VAST estimates that resulted in a higher error than the stratified mean were for the Fall season. These 15 Fall estimates span all other scenario variations and had a common theme of producing abundance estimates that are well above the true model value. Including detailed covariate information made a noticeable difference in our Yellowtail Flounder VAST estimates. Of the VAST estimates that included covariates information, $47/48\approx98$\% had a lower relative error than the corresponding stratified mean estimate. 
**For discussion: clearly our perfect covariate information made a big difference here, but the question is could we achieve similar results without "perfect" information? Which covariate is best? Good question for future research**

@@

As seen in Figure \@ref(fig:PopPct), in the constant temperature scenario the Cod population is decreasing its presence in stratas 19 and 20 over the duration of the simulation in both seasons, while simultaneously increasing their presence in strata 16. However, an increase in average temperature had a seasonal impact on Cod's presence in strata 16 in our simulations. In the spring the population followed a similar trend of leaving 19 and 20 while simultaneously entering 16. However, in the fall the increasing temperature drives the population out of strata 16 and 21, and into 18, 22, and 24. **check the impact of this difference on estiamtes**

>>>>

```{r strata-plot, echo=FALSE,out.width="95%",fig.show='hold', fig.align='center', fig.cap="Strata inhabited by each species in our population models."}
#knitr::include_graphics(c("Images/Strata.png"))
#knitr::include_graphics(c("Images/Excluded_strata.png"))
knitr::include_graphics(c("Images/Strata2.png"))
knitr::include_graphics(c("Images/Excluded_strata2.png"))
```

@@

Figure \@ref(fig:PopPct) depicts the spatial shifting that occurs in each strata within our population models, specifically during the bottom trawl survey weeks in the spring (13 and 14) and fall (37 and 28). See Figure \@ref(fig:strata-plot) for a spatial reference of the Georges Bank strata and Figure \@ref(fig:hab-plot) for the static habitat preferences for each species ($Hab^2_{J,s}$) used in our population models.

@@

The top two panels in Figure \@ref(fig:PopPct) depict the results for Yellowtail Flounder with a repeating temperature gradient on the left (constant temperature) and a temperature gradient that increases over time on the right (increasing temperature). In both temperature scenarios we see the percent of Yellowtail Flounder in strata 13 decrease over the course of the time series in both seasons. The spring population in stratum 19 also decreases in both temperature scenarios as well. The percent of the population increases in stratum 16 in the spring over the duration of both time series, which implies the flow out of strata 13 and 19 in the spring are going into stratum 16. These dynamics occur in both temperature scenarios in weeks 13 and 14 because stratum 16 contains favorable habitat for Yellowtail Flounder that coincides with most of the areas we have designated as the species' spawning ground, which takes place in weeks 9-12. These spring dynamics are therefore related to the static habitat values in our model rather than the dynamics temperature preferences. While we observe similar dynamics in the fall (weeks 37 and 38) in the constant temperature scenario, an increasing average temperature results in a decrease in population in stratum 16 over time and corresponding increases in strata 17 and 18 (see Figure \@ref(fig:PopPct)). These dynamics imply that an increase in temperature results in the more shallow strata 16 becoming less desirable than the deeper and more narrow exterior strata 17 and 18. One noticeable seasonal difference in the constant temperature scenario for Yellowtail is how ~10\% of the population exists in the narrow exterior strata 17 in the fall, while seemingly none of the population exists in any of the exterior strata (14, 15, 17, 18) in the spring.

Tables \@ref(tab:YTresultsallstrata) and \@ref(tab:YTresultsreducedstrata) contain the absolute error between our abundance estimates for Yellowtail Flounder comparing design-based approach (stratified mean) with two settings for a model-based approach (VAST A \& B). Of all scenarios without covariates, $33/48\approx68$\% had a VAST fit with a lower relative error compared to the stratified mean estimate. All 15 of the covariate-free VAST estimates that resulted in a higher error than the stratified mean were for the fall season and had a common theme of producing abundance estimates that are above the true model value. These 15 fall estimates span all other scenario variations. The implication of this is that our model-based approach without covariates struggles with the primary seasonal difference for Yellowtail Flounder, which is that a larger percentage of the population exists in the narrow exterior strata (18 and/or 17). This theory is further supported by the fact that the absolute error in the increasing temperature scenarios increased dramatically in the fall season, when the combined percentage of the population in the outer strata 17 and 18 increased to over 40\% by the end of the simulation.

Including detailed covariate information made a noticeable difference in our Yellowtail Flounder model-based VAST estimates. All of the VAST estimates that included covariate information produced a lower relative error than the corresponding VAST model that did not include covariates. The largest improvements were seen in the increasing temperature scenarios. As a result, of the VAST estimates that included covariates information, $47/48\approx98$\% had a lower relative error than the corresponding stratified mean estimate. This implies that the covariate information helped our model-based estimate account for the design-based issues related to small strata that contain a large percentage of the population, which can become exacerbated in the increasing temperature scenario.

The performance of estimates tend to decrease under increasing temperature, with the most dramatic changes seen in our VAST estimates without covariates. One exception to this is when we use a stratified mean approach while the population is decreasing. Since our analyses have found that the stratified mean tends to under estimate the true abundance, as the Yellowtail population decreases towards zero the difference between the estimate and the true value also decrease. That is, if the population is low enough, failing to appropriately sample the population produces the same result as appropriately sampling.

**For discussion: clearly our perfect covariate information made a big difference here, but the question is could we achieve similar results without "perfect" information? Which covariate is best? Good question for future research**

@@

As seen in Figure \@ref(fig:PopPct), in the constant temperature scenario the Cod population is decreasing its presence in strata 19 and 20 over the duration of the simulation in both seasons, while simultaneously increasing their presence in strata 16. However, an increase in average temperature had a seasonal impact on Cod's presence in strata 16 in our simulations. In the spring the population followed a similar trend of leaving 19 and 20 while simultaneously entering 16. However, in the fall the increasing temperature drives the population out of strata 16 and 21, and into 18, 22, and 24. **check the impact of this difference on estimates**

### 2023-03-06 Manuscript/Draft1.tex
update manuscript

{\centering \includegraphics[width=0.95\linewidth]{Images/Strata} \includegraphics[width=0.95\linewidth]{Images/Excluded_strata}

@@

@@

Figure \ref{fig:PopPct} depicts the spatial shifting that occurs in our population models in each of the survey weeks.

Of the scenarios without covariates, \(33/48\approx68\)\% had a VAST fit with a lower relative error compared to the stratified mean estimate. All 15 of the covariate-free VAST estimates that resulted in a higher error than the stratified mean were for the Fall season. These 15 Fall estimates span all other scenario variations and had a common theme of producing abundance estimates that are well above the true model value. Including detailed covariate information made a noticeable difference in our Yellowtail Flounder VAST estimates. Of the VAST estimates that included covariates information, \(47/48\approx98\)\% had a lower relative error than the corresponding stratified mean estimate.
\textbf{For discussion: clearly our perfect covariate information made a big difference here, but the question is could we achieve similar results without ``perfect'' information? Which covariate is best? Good question for future research}

@@

As seen in Figure \ref{fig:PopPct}, in the constant temperature scenario the Cod population is decreasing its presence in stratas 19 and 20 over the duration of the simulation in both seasons, while simultaneously increasing their presence in strata 16. However, an increase in average temperature had a seasonal impact on Cod's presence in strata 16 in our simulations. In the spring the population followed a similar trend of leaving 19 and 20 while simultaneously entering 16. However, in the fall the increasing temperature drives the population out of strata 16 and 21, and into 18, 22, and 24. \textbf{check the impact of this difference on estiamtes}

>>>>

{\centering \includegraphics[width=0.95\linewidth]{Images/Strata2} \includegraphics[width=0.95\linewidth]{Images/Excluded_strata2}

@@

Figure \ref{fig:PopPct} depicts the spatial shifting that occurs in each strata within our population models, specifically during the bottom trawl survey weeks in the spring (13 and 14) and fall (37 and 28). See Figure \ref{fig:strata-plot} for a spatial reference of the Georges Bank strata and Figure \ref{fig:hab-plot} for the static habitat preferences for each species (\(Hab^2_{J,s}\)) used in our population models.

@@

The top two panels in Figure \ref{fig:PopPct} depict the results for Yellowtail Flounder with a repeating temperature gradient on the left (constant temperature) and a temperature gradient that increases over time on the right (increasing temperature). In both temperature scenarios we see the percent of Yellowtail Flounder in strata 13 decrease over the course of the time series in both seasons. The spring population in stratum 19 also decreases in both temperature scenarios as well. The percent of the population increases in stratum 16 in the spring over the duration of both time series, which implies the flow out of strata 13 and 19 in the spring are going into stratum 16. These dynamics occur in both temperature scenarios in weeks 13 and 14 because stratum 16 contains favorable habitat for Yellowtail Flounder that coincides with most of the areas we have designated as the species' spawning ground, which takes place in weeks 9-12. These spring dynamics are therefore related to the static habitat values in our model rather than the dynamics temperature preferences. While we observe similar dynamics in the fall (weeks 37 and 38) in the constant temperature scenario, an increasing average temperature results in a decrease in population in stratum 16 over time and corresponding increases in strata 17 and 18 (see Figure \ref{fig:PopPct}). These dynamics imply that an increase in temperature results in the more shallow strata 16 becoming less desirable than the deeper and more narrow exterior strata 17 and 18. One noticeable seasonal difference in the constant temperature scenario for Yellowtail is how \textasciitilde10\% of the population exists in the narrow exterior strata 17 in the fall, while seemingly none of the population exists in any of the exterior strata (14, 15, 17, 18) in the spring.

Tables \ref{tab:YTresultsallstrata} and \ref{tab:YTresultsreducedstrata} contain the absolute error between our abundance estimates for Yellowtail Flounder comparing design-based approach (stratified mean) with two settings for a model-based approach (VAST A \& B). Of all scenarios without covariates, \(33/48\approx68\)\% had a VAST fit with a lower relative error compared to the stratified mean estimate. All 15 of the covariate-free VAST estimates that resulted in a higher error than the stratified mean were for the fall season and had a common theme of producing abundance estimates that are above the true model value. These 15 fall estimates span all other scenario variations. The implication of this is that our model-based approach without covariates struggles with the primary seasonal difference for Yellowtail Flounder, which is that a larger percentage of the population exists in the narrow exterior strata (18 and/or 17). This theory is further supported by the fact that the absolute error in the increasing temperature scenarios increased dramatically in the fall season, when the combined percentage of the population in the outer strata 17 and 18 increased to over 40\% by the end of the simulation.

Including detailed covariate information made a noticeable difference in our Yellowtail Flounder model-based VAST estimates. All of the VAST estimates that included covariate information produced a lower relative error than the corresponding VAST model that did not include covariates. The largest improvements were seen in the increasing temperature scenarios. As a result, of the VAST estimates that included covariates information, \(47/48\approx98\)\% had a lower relative error than the corresponding stratified mean estimate. This implies that the covariate information helped our model-based estimate account for the design-based issues related to small strata that contain a large percentage of the population, which can become exacerbated in the increasing temperature scenario.

The performance of estimates tend to decrease under increasing temperature, with the most dramatic changes seen in our VAST estimates without covariates. One exception to this is when we use a stratified mean approach while the population is decreasing. Since our analyses have found that the stratified mean tends to under estimate the true abundance, as the Yellowtail population decreases towards zero the difference between the estimate and the true value also decrease. That is, if the population is low enough, failing to appropriately sample the population produces the same result as appropriately sampling.

\textbf{For discussion: clearly our perfect covariate information made a big difference here, but the question is could we achieve similar results without ``perfect'' information? Which covariate is best? Good question for future research}

@@

As seen in Figure \ref{fig:PopPct}, in the constant temperature scenario the Cod population is decreasing its presence in strata 19 and 20 over the duration of the simulation in both seasons, while simultaneously increasing their presence in strata 16. However, an increase in average temperature had a seasonal impact on Cod's presence in strata 16 in our simulations. In the spring the population followed a similar trend of leaving 19 and 20 while simultaneously entering 16. However, in the fall the increasing temperature drives the population out of strata 16 and 21, and into 18, 22, and 24. \textbf{check the impact of this difference on estimates}
